### 2021-08-13 analysis/preliminaryAnalysis.Rmd
added some explanatory notes

---
title: "An assessment of statistical guidance for authors at influential academic journals across scientific disciplines: Preliminary analysis"
author: "Maia & Tom"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = T)
```

```{r load_packages}
library(knitr) # for literate programming
library(papaja) # for article template
library(tidyverse) # for data munging
library(tidylog) # inline analysis feedback
library(here) # for finding files
library(gtsummary) # for summary tables
```

```{r load_functions}
# load custom functions
source(here('analysis','functions.R'))
```

```{r perform_preprocessing}
# loads raw data, performs preprocessing, saves preprocessed files
source(here('analysis','preprocessing.R'))
```

```{r load_processed_data}
# loads the processed data files output by the preprocessing performed above
load(here('data','processed','d_coding.rds'))
load(here('data','processed','d_journals.rds'))
load(here('data','processed','lookup_external_guidance.rds'))
```

```{r set_style_defaults}
theme_gtsummary_compact() # set table theme
```

<br>
We have now completed data collection for the first phase of this project in which we extracted author-facing statistical guidance provided by the top-15 journals (by Impact Factor) in each of 22 scientific disciplines (i.e., total N = 330 journals). The preregistered study protocol is available [here](https://osf.io/cz9g3/).

> *Operational definition of statistical guidance*: guidance related to the appropriate selection, implementation, reporting, and interpretation of statistical analyses

We recorded whether each journal:

* Provided any statistical guidance at all
* Had a dedicated statistical guidance section
* Referred to statistical guidance in external sources (e.g., reporting guidelines or articles)

*Note:* When journals referred to guidance in external sources, we have recorded the names of the sources, but *we have currently not extracted guidance from those sources*. We return to this point in the *issues to address* section at the end of this document.

* If any statistical guidance was provided by the journal, we coded (YES/NO) whether it mentioned each of 20 statistical topics (e.g., "p-values"), and extracted the verbatim text.

All data extraction was performed by a first coder (DS, MM, TB, MSH) and a second coder (TEH) with differences resolved through discussion.

*Note* Some publishers provided statistical guidance that was shared across their portfolio journals (specifically, 32 Nature journals, 12 Cell journals, 2 Frontiers journals) . Publisher guidance was coded by two team members (as above), but is represented in our data and analysis for each individual journal the guidance applies to. For example, the Nature guidance is inherited by 32 journals so appears 32 times in the dataset.

Below we present a preliminary analysis of the phase one data.

Look at counts and proportions mentioning each topic. Use all journals for denominator. Overall and by field.

``````{r tbl_counts_props}
d_journals %>% 
  select(journal, esi_field) %>% 
  left_join(d_coding, ., by = "journal") %>%

  # Recode NAs to FALSE
  mutate(across(
    everything()
    # starts_with("has_") & (!contains("guidance")) # for internal guidance elements
    , ~replace_na(., FALSE) )) %>%

  select(esi_field, starts_with("has_")) %>% 
  gtsummary::tbl_summary(
    by = esi_field,
      # statistic = list(everything() ~ "{p}% ({n} / {N})")
  ) %>%
  add_overall %>% 
  modify_footnote(everything() ~ NA) %>% 
  modify_caption("**Number and proportion of journals overall and by field providing internal statistical guidance by topic.** Denominator for all proportions is the number of journals in the field, regardless of whether the journal has internal guidance.")
```

Some initial observations
- Math and computer science are indeed at 0% across the board.
- Clinical medicine has 100% guidance. Also high in related field: biology & biochemistry, immunology, molecular bio & genetics. Also psych and neuro fields.

Try a mosaic plot for internal guidance...nope!

```{r plot_internal_guidance_mosaic}
my_palette <- c("lightcoral", "lightgrey", "lightblue")

d_internal_guidance <-
  d_coding %>% 
  select(journal, starts_with("has_") & (!contains("guidance"))) %>% 
  pivot_longer(
    -journal,
    names_to = "topic",
    values_to = "outcome",
    values_drop_na = FALSE
  ) %>% 
  mutate(outcome = case_when(
    outcome ~ "Mentioned",
    !outcome ~ "Not mentioned",
    TRUE ~ "Not applicable (No internal guidance)"
  ))

# plt <- 
  d_internal_guidance %>%
  ggplot(aes(x = topic, y = factor(journal))) +
  geom_tile(aes(fill = outcome), colour = 'black', size = .25) +
  scale_fill_manual(
    values = my_palette,
    # labels = c('Not mentioned', 'Unclear', 'Mentioned')
  ) +
  scale_x_discrete(position = 'top') +
  papaja::theme_apa() +
  xlab("Topics mentioned") +
  theme(plot.margin = margin(0, 50, 0, 0),
        axis.ticks = element_blank(),
        axis.title.y =element_blank(),
        panel.grid.major=element_line(colour=NA), 
        panel.grid.minor=element_line(colour=NA),
        legend.position = 'bottom',
        axis.text.x  = element_text(angle = 45, hjust = 0),
        strip.background = element_rect(
          color="black", fill="white", size=0, linetype="solid"
        ),
        strip.text.y = element_text(
          size = 8, color = "black", face = 'bold'
        ))

# ggsave('internal-guidance.png', width = 11, height = 10)
```

Try bar chart

```{r plot_internal_guidance_bar}
d_internal_guidance %>% 
  ggplot(aes(y = topic, fill = outcome)) +
  geom_bar() +
  scale_fill_manual(values = my_palette)
```

Look at external guidance

```{r tbl_external_guidance}
lookup_external_guidance %>% 
  knitr::kable()
```

External guidance types

```{r tbl_external_guidance_types}
lookup_external_guidance %>% 
  count(guidance_type) %>% 
  kable()
```