### 2021-09-02 analysis/preliminaryAnalysis.Rmd
preliminary analysis 1st full draft

---
title: "Cross-disciplinary assessment of statistical guidance provided to authors by influential academic journals"
description: |
  Preliminary analysis for study one and study two.
author:
  - name: Maia Salholz-Hillel & Tom Hardwicke
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = T, fig.topcaption = TRUE)
```

```{r load_packages}
library(knitr) # for literate programming
library(tidyverse) # for data munging
library(janitor) # for data munging
library(tidylog) # inline code feedback
library(here) # for finding files
library(assertr) # for testing
library(lemon) # to have axes across all graph facets
library(ggthemr) # for ggplot theme
library(ggrepel) # for ggplot text labels
library(scales) # to wrap text on axis labels
library(rmarkdown) # for paged table
```

```{r load_functions}
source(here('analysis','functions.R')) # load custom functions
```

```{r perform_preprocessing}
# loads raw data, performs preprocessing, saves processed data files
#source(here('analysis','preprocessing.R'))
```

```{r load_processed_data}
# loads the processed data files
load(here('data','processed','d_all.rds'))
load(here('data','processed','lookup_external_guidance.rds'))
```

```{r set_defaults}
ggthemr('fresh') # set ggplot theme
```

# Introduction

We have now completed dual-coded data collection for the first study of this project and single-coded data collection for the second study of this project. The preregistered study protocol is available [here](https://osf.io/cz9g3/). Below we present a preliminary analysis of the data and highlight some issues to discuss as a team.

# Methods summary

## Study one

In November, 2019, we made copies of any documentation (e.g., instructions to authors) that provided author-facing statistical guidance provided by the top-15 journals (ranked by Impact Factor) in each of 22 scientific disciplines (i.e., total N = 330 journals).

**Operational definition of statistical guidance**

> any advice or instruction related to the appropriate selection, implementation, reporting, or interpretation of statistical analyses.

We then examined this documentation and recorded whether each journal:

* Provided any statistical guidance at all
* Had a dedicated statistical guidance section
* Referred to statistical guidance in external sources (e.g., reporting guidelines or articles)

If any statistical guidance was provided by a journal:

* We also recorded whether it mentioned each of 20 statistical topics (e.g., "p-values"), and extracted the verbatim text.

All data extraction was performed by a first coder (DS, MM, TB, MSH) and a second coder (TEH) with any coding differences resolved through discussion.

**Note:** When journals referred to guidance in external sources (typically reporting guidelines), we have recorded the names of the sources, but *we have not extracted guidance from those sources*.

```{r shared_publisher_journals}
# count number of journals that have shared publisher guidance
publisher_nature_n <- d_all %>% filter(publisher_guidance == "Nature") %>% nrow()
publisher_cell_n <- d_all %>% filter(publisher_guidance == "Cell") %>% nrow()
publisher_frontiers_n <- d_all %>% filter(publisher_guidance == "Frontiers") %>% nrow()
```

**Note:** We found that some publishers provided statistical guidance that was shared across their portfolio of journals (specifically, `r publisher_nature_n` Nature journals, `r publisher_cell_n` Cell journals, and `r publisher_frontiers_n` Frontiers journals). An additional journal - Scientific Data - shares guidance with other Nature journals, but also has its own specific guidance. Shared publisher guidance was coded by two team members (as above), but is represented in our data and analysis multiple times - once for each individual journal it applies to. For example, the `r publisher_cell_n` journals published by Cell inherit Cell's "STAR Methods" guidance, so this guidance appears `r publisher_cell_n` times in our dataset.

## Study two

For study two, we planned to more closely examine the guidance provided on five specific topics:

- p-values
- statistical significance
- confidence intervals
- effect sizes
- sample size justification

We prespecified some information we wanted to extract about these topics; however, when I (Tom) examined the verbatim guidance, I found our prespecified questions and response options to be unsuitable (ambiguous, badly worded, etc). So in an entirely data-dependent and exploratory manner, I instead coded the level of endorsement for each topic/method according to four categories:

1. Explicit endorsement - the journal advises or instructs authors to use this method whenever possible/appropriate.
2. Implicit endorsement - the journal provides advice on the method, implying endorsement, but does not explictly advise that the method should be used.
3. Implicit opposition - the journal advises that they would prefer the method is not used, but does not explicitly rule it out.
4. Explicit opposition - the journal advises that the method should not be used.

This was admittedly quite subjective and I had to 'read between the lines' a bit in order to assign guidance to these categories.

I also categorised reporting advice for three topics as below:

p-values:

1. Report exact p-values
2. Report exact p-values unless very small
3. Report p-values relative to thresholds (e.g., p < .05)
4. No guidance

Statistical significance:

1. Report alpha
2. Use specific alpha (e.g., a = .05)
3. No guidance

Confidence intervals:

1. Report confidence level
2. Use specific confidence level (e.g., 95%)
3. No guidance

So far, the data coding for study two has only been performed by me (Tom), though we originally planned for dual-coding (see issues section at the end of this document).

# Results

## Study one

```{r summarise_data}
sum_tab <- d_all %>%

  # Recode NAs to FALSE
  mutate(across(
    everything()
    , ~replace_na(., FALSE) )) %>% 
  select(esi_field, starts_with("has_")) %>% 
  group_by(esi_field) %>%
  summarise(across(starts_with("has_"), list(n = sum, N = ~n()), .names = "{col}__{fn}")) %>%
  adorn_totals(name = 'OVERALL') %>%
  pivot_longer(cols = !esi_field, names_to = c('variable', '.value'), names_sep = '__', names_prefix = 'has_') %>%
  mutate(prop = n/N, percent = round(prop*100, 0)) %>%
  mutate(esi_abbr = case_when( # create abbreviated field labels to use in figures
    esi_field == "OVERALL" ~ "OVERALL",
    esi_field == "AGRICULTURAL SCIENCES" ~ "AGRI",
    esi_field == "BIOLOGY & BIOCHEMISTRY" ~ "BIO",       
    esi_field == "CHEMISTRY" ~ "CHEM" ,                   
    esi_field == "CLINICAL MEDICINE" ~ "MED",           
    esi_field == "COMPUTER SCIENCE" ~ "COMSCI",             
    esi_field == "ECONOMICS & BUSINESS" ~ "ECON",         
    esi_field == "ENGINEERING" ~ "ENGIN",                  
    esi_field == "ENVIRONMENT_ECOLOGY" ~ "ECO",         
    esi_field == "GEOSCIENCES" ~ "GEO",                  
    esi_field == "IMMUNOLOGY" ~ "IMMUN",                   
    esi_field == "MATERIALS SCIENCE" ~ "MATSCI",            
    esi_field == "MATHEMATICS" ~ "MATH",                
    esi_field == "MICROBIOLOGY" ~ "MICBIO",                 
    esi_field == "MOLECULAR BIOLOGY & GENETICS" ~ "MOLBIO", 
    esi_field == "Multidisciplinary" ~ "MULTI",            
    esi_field == "NEUROSCIENCE & BEHAVIOR" ~ "NEURO",     
    esi_field == "PHARMACOLOGY & TOXICOLOGY" ~ "PHARM",    
    esi_field == "PHYSICS" ~ "PHYS",                     
    esi_field == "PLANT & ANIMAL SCIENCE" ~ "PLANT",       
    esi_field == "PSYCHIATRY_PSYCHOLOGY" ~ "PSY",       
    esi_field == "SOCIAL SCIENCES, GENERAL" ~ "SOCSCI",     
    esi_field == "SPACE SCIENCE" ~ "SPACE"          
  )) %>%
  mutate(esi_field_with_abbr = paste0(esi_field,' (', esi_abbr,')')) %>%
  mutate(var_display = case_when( # create display versions of topic names
    variable == "guidance" ~ "Provides any statistical guidance",
    variable == "internal_guidance" ~ "Provides journal-specific statistical guidance",
    variable == "internal_guidance_section" ~ "Has a dedicated statistical guidance section",
    variable == "external_guidance" ~ "Refers to external statistical guidance",
    variable == "p_value" ~ "p values",
    variable == "significance" ~ "statistical significance",
    variable == "null_hypo" ~ "null hypotheses",
    variable == "sample_size" ~ "sample size justification",
    variable == "conf_int" ~ "confidence intervals",
    variable == "effect_size" ~ "effect sizes",
    variable == "multi_compare" ~ "multiple comparisons",
    variable == "subgroup" ~ "subgroup analyses",
    variable == "baseline_covar" ~ "baseline covariates",
    variable == "non_param" ~ "non-parametric tests",
    variable == "sensitivity" ~ "sensitivity analyses",
    variable == "model_assume" ~ "checking model assumptions",
    variable == "exclusion" ~ "data exclusions",
    variable == "outliers" ~ "handling outliers",
    variable == "missing" ~ "handling missing data",
    variable == "one_sided" ~ "one-sided tests",
    variable == "bayes" ~ "Bayesian statistics",
    variable == "secondary" ~ "secondary outcomes",
    variable == "prespecify" ~ "prespecification of analyses",
    variable == "cat_continuous" ~ "categorisation of continuous data",
    variable == "publisher_guidance" ~ "Shares publisher guidance"
  ))

# get fields with no statistical guidance at all
null_fields <- sum_tab %>% 
  filter(
    variable == "guidance",
    prop == 0
  ) %>%
  pull(esi_field)

# select only variables that address the general any/internal/external statistical guidance classifications 
sum_tab_general <- sum_tab %>%
  filter(variable %in% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')) %>% # select only data on general classifications
  mutate(variable = factor(variable, levels = c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')),
         var_display = factor(var_display, levels = c("Provides any statistical guidance", "Provides journal-specific statistical guidance", "Has a dedicated statistical guidance section", "Refers to external statistical guidance"))) # set variable order

# select only variables on specific topics
sum_tab_topics <- sum_tab %>%
  filter(esi_field %notin% null_fields) %>% # remove fields that have no statistical guidance at all
  filter(variable %notin% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance', 'publisher_guidance')) %>% # select only data on topics
  rename(topic = variable, topic_display = var_display)

# get list of statistical topics in order of overall proportions
ordered_topics <- sum_tab_topics %>%
  filter(esi_field == "OVERALL") %>%
  arrange(desc(prop)) %>% # order from highest to lowest
  pull(topic_display)

# set order of topics
sum_tab_topics <- sum_tab_topics %>%
  mutate(topic_display = factor(topic_display, levels = ordered_topics))
```

Figure \@ref(fig:statguidefig) shows the percentage of journals in each scientific discipline (denominator is N = 15) and overall (denominator is N = 330) that:

(a) offered any statistical guidance; 
(b) had a dedicated statistical guidance section.

For tabular data, see the table in the Appendix.

```{r crosstab_int_ext}
crosstab_int_ext <- d_all %>% count(has_internal_guidance, has_external_guidance)
only_ext <- crosstab_int_ext %>% filter(has_internal_guidance == F, has_external_guidance == T) %>% pull(n)
```

You can see that around half (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "guidance") %>% pull(percent)`%) of journals offered any statistical guidance. Note that this includes `r only_ext` journals which only referred to statistical guidance in external sources (reporting guidelines or academic papers). Just over quarter of journals had a dedicated statistical guidance section in their author instructions (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "internal_guidance_section") %>% pull(percent)`%). In two fields (Computer Science and Maths), no journals offered any statistical guidance. Journals in health-related fields were more likely to offer statistical guidance and have dedicated statistical guidance sections. Notably, 100% of the journals in clinical medicine offered some statistical guidance.

```{r}
# extract overall data
sum_tab_general_overall <- sum_tab_general %>%
  filter(variable %in% c('guidance', 'internal_guidance_section')) %>%
  filter(esi_field == "OVERALL")
```

```{r statguidefig, layout="l-page", fig.width = 10, fig.height = 6, fig.cap = "Percentage of journals offering statistical guidance by scientific discipline (N = 15; represented by coloured dots) and overall (N = 330; represented by black diamond)."}
sum_tab_general %>%
  filter(esi_field != "OVERALL") %>%
  filter(variable %in% c('guidance', 'internal_guidance_section')) %>%
  ggplot(aes(x=percent, y = 1, colour = str_wrap(esi_field_with_abbr,25))) +
  facet_rep_wrap(facets = vars(var_display), ncol = 2, nrow = 1) +
  coord_capped_cart(bottom='both', left='both') + # to break axes in bottom left corner
  geom_point(size = 5, position = position_stack(), alpha = .5) +
  geom_point(size = 6, position = position_stack(), data = sum_tab_general_overall, shape = 18, colour = 'black', aes(fill = 'OVERALL')) +
  geom_text_repel(
    aes(label = ifelse(percent > 0,esi_abbr,''),
        segment.square = T,
        segment.inflect = T), #
    #point.padding = 0.2,
    force_pull   = 0, # do not pull toward data points
    nudge_y      = 15,
    #nudge_x = 5,
    direction    = "x",
    angle        = 90,
    hjust        = 0,
    segment.size = 0.2,
    segment.curvature = -1e-20,
    max.iter = 1e4, max.time = 1,
    max.overlaps = Inf, # ensure all labels are shown even if overlapping
    show.legend = FALSE # do not add legend entry
  ) +
  scale_y_continuous(limits = c(0,15), name = '1 dot per scientific discipline') +
  xlim(0,100) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = 'bottom') +
  guides(colour = guide_legend(nrow = 8, title.position = 'top', title = '')) +
  guides(fill = guide_legend(nrow = 1, title.position = 'top', title = "Scientific disciplines"))
```
Figure \@ref(fig:stattopicsfig) shows the percentage of journals in each scientific discipline (denominator is N = 15) and overall (denominator is N = 330) that mentioned each of the twenty statistical topics we pre-selected. For tabular data, see the table in the Appendix. Please see the issues section at the end of this document for an important note regarding the topic "prespecification of analyses".

You can see that some topics (e.g., confidence intervals, p-values) are much more likely to be mentioned than other topics (e.g., handling outliers, categorisation of continuous data). Journals in health-related fields, particularly clinical medicine, are often more likely to mention this collection of topics than journals in other fields.

```{r stattopicsfig, layout="l-page", fig.width = 10, fig.height = 28, fig.cap = "Percentage of journals offering guidance on twenty statistical topics by scientific discipline (N = 15; represented by coloured dots) and overall (N = 330; represented by black diamond). For presentational purposes, two scientific disciplines in which no journals offered any statistical guidance at all are not shown. Additionally, disciplines offering no guidance on individual topics are shown, but not labelled. Graphs are ordered from left to right and top to bottom in order of highest proportion offering topic guidance overall across scientific disciplines."}
# extract overall data
sum_tab_topics_overall <- sum_tab_topics %>%
  filter(esi_field == "OVERALL")

sum_tab_topics %>%
  filter(esi_field != "OVERALL") %>%
  ggplot(aes(x=percent, y = 1, colour = str_wrap(esi_field_with_abbr,25))) +
  facet_rep_wrap(facets = vars(topic_display), ncol = 2) +
  coord_capped_cart(bottom='both', left='both') + # to break axes in bottom left corner
  geom_point(size = 5, position = position_stack(), alpha = .5) +
  geom_point(size = 6, position = position_stack(), data = sum_tab_topics_overall, shape = 18, colour = 'black', aes(fill = 'OVERALL')) +
  geom_text_repel(
    aes(label = ifelse(percent > 0,esi_abbr,''),
        segment.square = T,
        segment.inflect = T), #
    #point.padding = 0.2,
    force_pull   = 0, # do not pull toward data points
    nudge_y      = 25,
    nudge_x = 10,
    direction    = "x",
    angle        = 90,
    hjust        = 0,
    segment.size = 0.2,
    segment.curvature = -1e-20,
    max.iter = 1e4, max.time = 1,
    max.overlaps = Inf, # ensure all labels are shown even if overlapping
    show.legend = FALSE # do not add legend entry
  ) +
  scale_y_continuous(limits = c(0,25), name = '1 dot per scientific discipline') +
  xlim(0,100) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = 'bottom') +
  guides(colour = guide_legend(order = 1, nrow = 8, title.position = 'top', title = '')) +
  guides(fill = guide_legend(order = -1, nrow = 1, title.position = 'top', title = "Scientific disciplines"))
```

```{r how_much_guidance}
n_topics <- d_all %>% filter(has_guidance == T, has_internal_guidance == T) %>%

  # Recode NAs to FALSE
  mutate(across(
    everything()
    , ~replace_na(., FALSE) )) %>% 
  select(journal, starts_with("has_"), -c(has_guidance, has_internal_guidance, has_internal_guidance_section, has_external_guidance, has_publisher_guidance)) %>%
  pivot_longer(cols = !journal, names_to = 'topic', values_to = 'value') %>%
  group_by(journal) %>%
  summarise(topics_mentioned = sum(value))
```

For the `r d_all %>% filter(has_guidance == T, has_internal_guidance == T) %>% nrow()` journals that offered some internal guidance (i.e., excluding those that offered no guidance at all and those that only referred to external sources), the histogram in Figure \@ref(fig:topicshist) illustrates that the maximum number of topics mentioned by an individual journal was `r max(n_topics$topics_mentioned)`. There were `r n_topics %>% filter(topics_mentioned == 0) %>% nrow()` journals that did not mention any of our prespecified topics (but provided statistical guidance on other topics). The median topics mentioned was `r median(n_topics$topics_mentioned)`.

```{r topicshist, fig.cap = 'Histogram showing how many of the twenty statistical topics were mentioned by each journal'}
n_topics %>%
  ggplot(aes(x=topics_mentioned)) +
  geom_histogram()
```
In the grid below (Figure \@ref(fig:guidancemosaic)), you can see which of the twenty preselected topics were mentioned by each journal. Journals that shared publisher guidance are represented by one row (per publisher). Journals that did not offer any statistical guidance at all, or journals that referred only to guidance in external sources, are not shown.

```{r}
my_palette <- c("#95D840FF", "#404788FF")

d_internal_guidance_noPublisher <-
  d_all %>%
  filter(has_internal_guidance == T, # select journals with internal guidance
         has_publisher_guidance == F | journal == "Scientific Data") # remove journals with shared publisher guidance except for Scientific Data (which shared guidance and its own guidance)

# were going to represent journals with publisher guidance in a single row per publisher. 
# select exemplar journals for each publisher
publisher_nature <- d_all %>% filter(journal == "Nature") %>% mutate(journal = paste0("Nature journals"," (n=", publisher_nature_n,')'))
publisher_cell <- d_all %>% filter(journal == "CELL") %>% mutate(journal = paste0("Cell journals"," (n=", publisher_cell_n,')'))
publisher_frontiers <- d_all %>% filter(journal == "Frontiers in Microbiology") %>% mutate(journal = paste0("Frontiers journals"," (n=", publisher_frontiers_n,')'))

# combine the journals with no publisher guidance with the publisher guidance rows
d_internal_guidance <- bind_rows(d_internal_guidance_noPublisher, publisher_nature, publisher_cell, publisher_frontiers)

# list of journals in order of most topics mentioned
journals_topics_mentioned_ordered <- d_internal_guidance %>%
  # Recode NAs to FALSE
  mutate(across(
    everything()
    , ~replace_na(., FALSE) )) %>% 
  select(journal, starts_with("has_"), -c(has_guidance, has_internal_guidance, has_internal_guidance_section, has_external_guidance, has_publisher_guidance)) %>%
  pivot_longer(cols = !journal, names_to = 'topic', values_to = 'value') %>%
  group_by(journal) %>%
  summarise(topics_mentioned = sum(value)) %>%
  arrange(topics_mentioned) %>%
  pull(journal)

plt <- d_internal_guidance %>%
  mutate(journal = factor(journal, levels = journals_topics_mentioned_ordered)) %>% # order journals by n topics mentioned
  select(journal, starts_with("has_") & (!contains("guidance"))) %>%
  mutate(journal = fct_recode(journal,"PNAS" = "PROCEEDINGS OF THE NATIONAL ACADEMY OF SCIENCES OF THE UNITED STATES OF AMERICA")) %>% # shorten a particularly long journal name
  pivot_longer(
    -journal,
    names_to = "topic",
    names_prefix = 'has_',
    values_to = "outcome",
    values_drop_na = FALSE
  ) %>%
  mutate(outcome = case_when(
    outcome ~ "Mentioned",
    !outcome ~ "Not mentioned",
    TRUE ~ "Not applicable (No internal guidance)"
  )) %>%
  mutate(topic_display = case_when( # create display versions of topic names
    topic == "guidance" ~ "Provides any statistical guidance",
    topic == "internal_guidance" ~ "Provides journal-specific statistical guidance",
    topic == "internal_guidance_section" ~ "Has a dedicated statistical guidance section",
    topic == "external_guidance" ~ "Refers to external statistical guidance",
    topic == "p_value" ~ "p values",
    topic == "significance" ~ "statistical significance",
    topic == "null_hypo" ~ "null hypotheses",
    topic == "sample_size" ~ "sample size justification",
    topic == "conf_int" ~ "confidence intervals",
    topic == "effect_size" ~ "effect sizes",
    topic == "multi_compare" ~ "multiple comparisons",
    topic == "subgroup" ~ "subgroup analyses",
    topic == "baseline_covar" ~ "baseline covariates",
    topic == "non_param" ~ "non-parametric tests",
    topic == "sensitivity" ~ "sensitivity analyses",
    topic == "model_assume" ~ "checking model assumptions",
    topic == "exclusion" ~ "data exclusions",
    topic == "outliers" ~ "handling outliers",
    topic == "missing" ~ "handling missing data",
    topic == "one_sided" ~ "one-sided tests",
    topic == "bayes" ~ "Bayesian statistics",
    topic == "secondary" ~ "secondary outcomes",
    topic == "prespecify" ~ "prespecification of analyses",
    topic == "cat_continuous" ~ "categorisation of continuous data",
    topic == "publisher_guidance" ~ "Shares publisher guidance"
  )) %>%
  mutate(topic_display = factor(topic_display, levels = ordered_topics)) %>% # order topics by n topics mentioned
  ggplot(aes(x = topic_display, y = journal)) +
    geom_tile(aes(fill = outcome), colour = 'black', size = .25, alpha = .6) +
    scale_fill_manual(
      values = my_palette,
      # labels = c('Not mentioned', 'Unclear', 'Mentioned')
    ) +
    scale_x_discrete(position = 'top') +
    scale_y_discrete(labels = wrap_format(25)) + # wrap text on y axis
    xlab("Statistical topics") +
    theme(plot.margin = margin(0, 50, 0, 0),
          axis.ticks = element_blank(),
          axis.title.y =element_blank(),
          axis.line = element_blank(),
          panel.grid.major=element_line(colour=NA),
          panel.grid.minor=element_line(colour=NA),
          legend.position = 'top',
          axis.text.x  = element_text(angle = 45, hjust = 0),
          strip.background = element_rect(
            color="black", fill="white", size=0, linetype="solid"
          ),
          strip.text.y = element_text(
            size = 8, color = "black", face = 'bold'
          ))

# ggsave('internal-guidance.png', width = 11, height = 10)
```

```{r guidancemosaic, layout="l-page", fig.cap = "Grid diagram showing whether each journal provided guidance on each of twenty preselected statistical topics. Topics are ordered from most (left) to least (right) mentioned. Journals are ordered from most (top) to least (bottom) preselected topics mentioned.", fig.height = 35, fig.width = 10}
plt
```

```{r guidance_examples}
# get the names of the twenty statistical topics
topic_names <- d_all %>% 
  select(starts_with("has_"), -c(has_guidance, has_external_guidance, has_internal_guidance, has_internal_guidance_section, has_publisher_guidance)) %>% 
  rename_all(~str_replace(.,"^has_","")) %>% # remove has_ prefix
  verify(ncol(.) == 20) %>% # check there are twenty columns (topics) as expected
  colnames()

set.seed(42)
guidance_examples_tbl <- d_all %>%
  select(journal, topic_names) %>% 
  verify(ncol(.) == 21) %>% # check there are 21 columns as expected (i.e., journal and twenty topics)
  pivot_longer(cols = !journal, names_to = 'topic', values_to = 'verbatim') %>%
  distinct(verbatim, .keep_all = T) %>%
  drop_na(verbatim) %>%
  filter(nchar(verbatim) < 800) %>% # select shorter guidance only
  group_by(topic) %>%
  sample_n(2) %>%
  ungroup() %>%
  mutate(verbatim = paste0(verbatim,' (',journal,')')) %>% # add journal name to verbatim column
  select(-journal) %>% # remove journal column
  mutate(topic_display = case_when( # create display versions of topic names
    topic == "p_value" ~ "p values",
    topic == "significance" ~ "statistical significance",
    topic == "null_hypo" ~ "null hypotheses",
    topic == "sample_size" ~ "sample size justification",
    topic == "conf_int" ~ "confidence intervals",
    topic == "effect_size" ~ "effect sizes",
    topic == "multi_compare" ~ "multiple comparisons",
    topic == "subgroup" ~ "subgroup analyses",
    topic == "baseline_covar" ~ "baseline covariates",
    topic == "non_param" ~ "non-parametric tests",
    topic == "sensitivity" ~ "sensitivity analyses",
    topic == "model_assume" ~ "checking model assumptions",
    topic == "exclusion" ~ "data exclusions",
    topic == "outliers" ~ "handling outliers",
    topic == "missing" ~ "handling missing data",
    topic == "one_sided" ~ "one-sided tests",
    topic == "bayes" ~ "Bayesian statistics",
    topic == "secondary" ~ "secondary outcomes",
    topic == "prespecify" ~ "prespecification of analyses",
    topic == "cat_continuous" ~ "categorisation of continuous data"
  )) %>%
  select(topic = topic_display,"example guidance" =verbatim)
```

To give you a flavour of the kind of guidance provided, I have randomly selected two examples (excluding a few very long examples), for each of the twenty statistical topics, and displayed them in Table \@ref(tab:guidanceexamples) below.

```{r guidanceexamples}
guidance_examples_tbl %>% kable(caption = "Randomly selected verbatim statistical guidance for each of 20 preselected statistical topics.")
```

`r d_all %>% filter(has_external_guidance == T) %>% nrow()` journals referred authors to statistical guidance in an external source (reporting guidelines or other sources like academic papers or websites). Of these, `r crosstab_int_ext %>% filter(has_external_guidance == T, has_internal_guidance == F) %>% pull(n)` only referred to statistical guidance in external sources, whereas the other `r crosstab_int_ext %>% filter(has_external_guidance == T, has_internal_guidance == T) %>% pull(n)` also provided their own guidance in additional to referring to external sources.

In the histogram (Figure \@ref(fig:repguidehist)) below, you can see how frequently journals referred to different reporting guidelines. Then in Table \@ref(tab:tbl-external-other-guidance) you can see all of the other external guidelines (academic papers and websites) that journals referred to. None of these external sources that have been examined by us in detail.

```{r repguidehist, layout="l-page", fig.width = 10, fig.cap = 'Histogram showing how frequently journals referred to different reporting guidelines.'}
ext_guide_ordered <- lookup_external_guidance %>% arrange(desc(n_journals)) %>% pull(external_guidance)
lookup_external_guidance %>%
  filter(guidance_type == "reporting guideline") %>%
  mutate(external_guidance = factor(str_to_upper(external_guidance), levels = str_to_upper(ext_guide_ordered)),
         external_guidance = fct_recode(external_guidance, 
                                        `CONSORT PRO` = "CONSORT PRO EXTENSION",
                                        `CONSORT Cluster` = "CONSORT EXTENDED GUIDELINES",
                                        `HuGENet` = "HUMAN GENOME EPIDEMIOLOGY NETWORK (HUGENET) GUIDELINES",
                                        `FDA Diagnostic Tests` = "FDA GUIDELINES (HTTPS://PERMA.CC/W2EY-MSTA)",
                                        `CONSORT abstract` = "CONSORT ABSTRACT EXTENSION",
         NIH = "NIH PRINCIPLES AND GUIDELINES FOR REPORTING PRECLINICAL RESEARCH",
         NLM = "NLM RESEARCH REPORTING GUIDELINES AND INITIATIVES (HTTPS://WWW.NLM.NIH.GOV/SERVICES/RESEARCH_REPORT_GUIDE.HTML)")) %>%
  filter(guidance_type == "reporting guideline") %>%
  ggplot(aes(x=external_guidance, y = n_journals)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylim(0,100) +
  ylab('number of journals') +
  xlab('reporting guideline')
```
Other external sources (not reporting guidelines):

```{r tbl-external-other-guidance}
lookup_external_guidance %>%
  filter(guidance_type != "reporting guideline") %>%
  select(-guidance_type) %>%
  kable(caption = "Frequency of other external sources of statistical guidance (not reporting guidelines)")
```

## Study Two

Study two involved a closer look at five preselected statistical topics. Table \@ref(tab:s2table) shows how many journals offerred different levels of endorsement for these statistical methods. You can see that it was fairly rare to oppose the use of any of these methods. A few journals are taking a stand against the use of "statistical significance", but this is rare. Few journals are strongly advocating the use of p-values, but the vast majority find them acceptable. It was quite common to recommend reporting of effect sizes and confidence intervals.

```{r}
d_s2 <- read_csv(here('data','primary','data_phase_two.csv'))
```

```{r s2table}
d_s2 %>% select(journal, esi_field, p_use, stat_sig_use, conf_int_use, effect_size_use, sample_size_justification) %>%
  pivot_longer(cols = p_use:sample_size_justification, names_to = "topic") %>%
  filter(value != "No guidance") %>%
  count(topic, value) %>%
  mutate(value = fct_recode(value, `Implicit opposition` = 'Implicit opposition (for low sample sizes)',
         `Explicit opposition` = 'Explicit opposition (without prespecified multiplicity corrections)')) %>%
  pivot_wider(id_cols = topic, names_from = value, values_from = n) %>%
  mutate(across(everything(), .fns = ~replace_na(.,0))) %>% 
  mutate(topic = fct_recode(topic,
                            `confidence intervals` = "conf_int_use",
                            `effect sizes` = "effect_size_use",
                            `p-values` = "p_use",
                            `sample size justification` = "sample_size_justification",
                            `statistical significance` = "stat_sig_use")) %>%
  kable(caption = "Number of journals offering degrees of endorsement for five specific statistical methods")
```

Note that the one journal with implicit opposition to confidence intervals said this was only in the case of small sample sizes. And the one journal with explicit opposition to p-values said this was only when there were no pre-specified multiplicity corrections.

Of the `r d_s2 %>% filter(!is.na(p_value)) %>% nrow()` journals offering guidance on p-values, `r d_s2 %>% filter(!is.na(p_value)) %>% count(p_reporting) %>% filter(p_reporting == "Exact") %>% pull(n)` advised reporting of exact p-values, `r d_s2 %>% filter(!is.na(p_value)) %>% count(p_reporting) %>% filter(p_reporting == "Exact unless very small") %>% pull(n)` advised reporting of exact p-values unless they were very small (e.g., < .001), `r d_s2 %>% filter(!is.na(p_value)) %>% count(p_reporting) %>% filter(p_reporting == "Thresholds [0.05, 0.01, 0.001]") %>% pull(n)` advised to use reporting thresholds (e.g., <.05, <.01, <.001), and `r d_s2 %>% filter(!is.na(p_value)) %>% count(p_reporting) %>% filter(p_reporting == "No guidance") %>% pull(n)` offered no guidance on how to report p-values.

Of the `r d_s2 %>% filter(!is.na(significance)) %>% nrow()` journals offering guidance on statistical significance, `r d_s2 %>% filter(!is.na(significance)) %>% count(stat_sig_reporting) %>% filter(stat_sig_reporting == "Report alpha") %>% pull(n)` advised authors report the alpha level they used, `r d_s2 %>% filter(!is.na(significance)) %>% count(stat_sig_reporting) %>% filter(stat_sig_reporting == "0.05") %>% pull(n)` advised using an alpha level of .05, `r d_s2 %>% filter(!is.na(significance)) %>% count(stat_sig_reporting) %>% filter(stat_sig_reporting == "0.01") %>% pull(n)` advised using an alpha level of .01, and `r d_s2 %>% filter(!is.na(significance)) %>% count(stat_sig_reporting) %>% filter(stat_sig_reporting == "No guidance") %>% pull(n)` offered no guidance on reporting statistical significance.

Of the `r d_s2 %>% filter(!is.na(conf_int)) %>% nrow()` journals offering guidance on confidence intervals, `r d_s2 %>% filter(!is.na(conf_int)) %>% count(conf_int_reporting) %>% filter(conf_int_reporting == "Report 95% Cis") %>% pull(n)` advised authors to report 95% confidence intervals, `r d_s2 %>% filter(!is.na(conf_int)) %>% count(conf_int_reporting) %>% filter(conf_int_reporting == "Report confidence level") %>% pull(n)` advised authors report their chosen confidence level, and `r d_s2 %>% filter(!is.na(conf_int)) %>% count(conf_int_reporting) %>% filter(conf_int_reporting == "No guidance") %>% pull(n)` offered no guidance on reporting confidence intervals.

# Issues to address {#issues}

* In two fields (Computer Science and Maths), no journals offered any statistical guidance - I do not know these fields well, but I wonder if this indicates that statistical tools are rarely used, in which case one might reasonably suggest we should not even have bothered to examine these fields. Any thoughts on this?

* Our coding for the topic "prespecification of analyses" was not consistent across coders. Specifically, some coders considered that this topic also covered any guidance/instructions related to registration of clinical trials, whilst others did not. I believe the justification of those who did not was that clinical trials registration does not necessarily involve prespecification of a statistical analysis plan. That seems a reasonable justification to me, but I'd be interested to hear thoughts from the team. If we decide that advice about clinical trials registration alone does not fall into the remit of this topic, then it is relatively straightforward to examine the verbatim text we extracted and re-code the relevant cases. However, if we decide that advice about clinical trials registration *does* fall under the remit of this topic, then we will need to re-visit the journal instructions to authors and search for such guidance (as this was not done consistently previously).

* We have recorded whether journals refer to statistical guidance from external sources, such as reporting guidelines or academic papers. However, we have not examined those external sources, and we need to decide whether to do so. If we perform extraction and coding for all `r lookup_external_guidance %>% nrow()` external sources we have identified, this is clearly a lot of extra work. A counter-argument is that we are only really interested in the most salient statistical guidance offered directly by journals. On the other hand, if we do not extract information from external sources, then perhaps we are missing an important aspect of statistical guidance i.e., that journals may not provide their own guidance on certain topics because they are aptly covered by external sources. An additional complication is that in some cases, the boundary of internal vs external guidance is blurred when the 'external' source is an academic article that appears to have been written by members of the journal's editorial team.

* In the protocol we say that for Study 2, data extraction and coding will be in duplicate. However, we do caveat this by saying it might be necessary to deviate from the plan depending on our resources. I have done all of the first coding, do we have the capacity to do duplicate coding? My worry is that it will slow us down considerably, and I am keen to get these results published before they are outdated (the statistical guidance was originally extracted in November, 2019).

# Appendix {#app}

Tabular data (represented in Figures 1 and 2). Number and percentage of journals overall and by scientific field providing internal statistical guidance by topic (click the black arrow to scroll through the data for each field). The denominator for proportions is the number of journals in the field (N = 15) or overall (N = 330).

```{r tbl-counts-props}
sum_tab %>%
  filter(variable != "publisher_guidance") %>%
  mutate(field = paste0(esi_field,' N = ',N),
         table_data = paste0(n,' (',percent,'%)')) %>%
  select(field,variable,table_data) %>%
  pivot_wider(id_cols = variable, names_from = field, values_from = table_data) %>%
  select(variable, `OVERALL N = 330`, everything()) %>%
  paged_table(options = list(rows.print = 25, cols.print = 5))
```

```{r phase_two}
# p2_done <- read_csv(here('data','raw','phase_two_pval.csv'))
# # export data for phase two
# p2 <- d_all %>% filter(has_p_value == T | has_significance == T | has_effect_size == T | has_conf_int == T | has_sample_size == T) %>% 
#   mutate(coder = "TEH") %>% 
#   select(journal, esi_field, p_value, significance, conf_int, effect_size, sample_size) %>%
#   mutate(conf_int_use = NA,
#          effect_size_use = NA,
#          effect_size_reporting = NA,
#          sample_size_justification = NA,
#          sample_size_power = NA) 
# 
# full_join(p2, p2_done, 
#           by = c('journal','esi_field', 'p_value', 'significance')) %>%
#   arrange(p_value, significance) %>%
#   write_csv(here('data','raw','templates','phase_two.csv'))
# 
# tmp %>% filter(!is.na(significance))
# d_all %>% filter(has_significance == T)
```